---
title: "Class09"
author: "Sharon Chang"
date: "4/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)
wisc.df
```

```{r}
wisc.data <- as.matrix(wisc.df[,3:32])
row.names(wisc.data) <- wisc.df$id
head(wisc.data)
```

> How many observations are in this dataset?

```{r}
nrow(wisc.df)
```

> How many variables/features in the data are suffixed with _mean?

```{r}
length(grep("_mean", colnames(wisc.data)))
#colnames(wisc.data)
```


```{r}
diagnosis <- as.numeric(wisc.df$diagnosis == "M")
diagnosis
```

> How many are malignant?

```{r}
table(wisc.df$diagnosis)
```

```{r}
# Check column means and standard deviations
round(colMeans(wisc.data),1)
apply(wisc.data,2,sd)
```

```{r}
# Perform PCA on wisc.data
wisc.pr <- prcomp(wisc.data, scale = TRUE)

# Look at summary of results
summary(wisc.pr)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=(diagnosis+1), xlab = "PC1", ylab = "PC3")
```

```{r}
summary(wisc.pr)
```

> From your results, what proportion of the original variance is captured by the first principal components (PC1)?
0.4427

> How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
3

> How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
7

```{r}
# Calculate variance of each component
pr.var <- wisc.pr$sdev^2
head(pr.var)
```

```{r}
# Variance explained by each principal component: pve
pve <- (pr.var / sum(pr.var))
head(pve)

par(mfcol=c(1,2))

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")

# Plot cumulative proportion of variance explained
plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```


```{r}
## ggplot based graph
library(factoextra)
```

```{r}
fviz_eig(wisc.pr, addlabels = TRUE)
```


```{r}
# Scale the wisc.data data: data.scaled
data.scaled <- scale(wisc.data)
```

```{r}
data.dist <- dist(data.scaled)
```

```{r}
wisc.hclust <- hclust(data.dist, method="complete")
```

```{r}
plot(wisc.hclust)
abline(h=19, col="red", lty=2)
```

```{r}
#Use cutree() to cut the tree so that it has 4 clusters. Assign the output to the variable wisc.hclust.clusters.

wisc.hclust.clusters <- cutree(wisc.hclust, k = 4)


#We can use the table() function to compare the cluster membership to the actual diagnoses.

table(wisc.hclust.clusters, diagnosis)
```

```{r}
wisc.pca.hclust <- hclust(dist(wisc.pr$x[,1:7]), method = "ward.D2")
plot(wisc.pca.hclust)
```

```{r}
grps <- cutree(wisc.pca.hclust, k=2)
table(grps)
table(grps, diagnosis)
```

```{r}
#table(, diagnosis)
table(wisc.hclust.clusters, diagnosis)
```


```{r}
plot(wisc.pr$x[,1:2], col=grps)
plot(wisc.pr$x[,1:2], col=diagnosis+1)
```

```{r}
library(rgl)
plot3d(wisc.pr$x[,1:3], xlab="PC 1", ylab="PC 2", zlab="PC 3", cex=1.5, size=1, type="s", col=diagnosis+1)
```


```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
points(npc[,1], npc[,2], col="blue", pch=16, cex = 2)
```


