---
title: "PCA"
author: "Sharon Chang"
date: "4/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
x <- read.csv("UK_foods.csv")

## Complete the following code to find out how many rows and columns are in x?
dim(x)
```

```{r}
## Preview the first 6 rows
head(x)
```

```{r}
# Note how the minus indexing works
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
```

```{r}
# Check the dimensions again
dim(x)
```

```{r}
x <- read.csv("UK_foods.csv", row.names=1)
head(x)
```

> Q2. Which approach to solving the ‘row-names problem’ mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?

> The second method (row.names = 1). You cannot run minus indexing multiple times (it will delete a column every time)

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

> Q3: Changing what optional argument in the above barplot() function results in the following plot?

> Setting beside = FALSE

> Q5: Generating all pairwise plots may help somewhat. Can you make sense of the following code and resulting figure? What does it mean if a given point lies on the diagonal for a given plot?

```{r}
pairs(x, col=rainbow(10), pch=16)
```

> Q6. What is the main differences between N. Ireland and the other countries of the UK in terms of this data-set?

```{r}
# Use the prcomp() PCA function 
pca <- prcomp( t(x) )
summary(pca)
```

```{r}
# Plot PC1 vs PC2
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500), col=rainbow(4))
text(pca$x[,1], pca$x[,2], colnames(x), col=rainbow(4))
```

```{r}
# Below we can use the square of pca$sdev , which stands for “standard deviation”, to calculate how much variation in the original data each PC accounts for.

v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
```

```{r}
## or the second row here...
z <- summary(pca)
z$importance
```

```{r}
# This information can be summarized in a plot of the variances (eigenvalues) with respect to the principal component number (eigenvector number), which is given below.

barplot(v, xlab="Principal Component", ylab="Percent Variation")
```

```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```

```{r}
# Another way to see this information together with the main PCA plot is in a so-called biplot:

## The inbuilt biplot() can be useful for small datasets 
biplot(pca)
```

