---
title: "Class08"
author: "Sharon Chang"
date: "4/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

Let's start with an example of running the **kmeans()** function
```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```

> Use the kmeans() function setting k to 2 and nstart=20

```{r}
km <- kmeans(x, centers = 2, nstart = 20)
km
```

> Inspect/print the results
> Q. How many points are in each cluster?
30

> Q. What ‘component’ of your result object details
  - cluster size?
 - cluster assignment/membership?
  - cluster center?
  
```{r}
km$size
km$center
```
  
> Plot x colored by the kmeans cluster assignment and
add cluster centers as blue points

```{r}
plot(x, col = km$cluster)
points(km$centers, pch = 18, col = "blue", cex = 3)
```

## Hierarchical Clustering Example

We must give the **hclust()** function a distsance matrix not the raw data as input

```{r}
# Distance matrix calculation
d <- dist

# Clustering
hc <- hclust(dist(x))
plot(hc)
abline(h = 6, col = "red")
# cutree(hc, h = 6) # Cut by height h
cutree(hc, k = 2) # Cut by k groups
```

# Step 1. Generate some example data for clustering
```{r}
x <- rbind(
matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
```

# Step 2. Plot the data without clustering
```{r}
plot(x)
```

# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
```{r}
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

# Your Turn!
Q. Use the dist(), hclust(), plot() and cutree()
functions to return 2 and 3 clusters
```{r}
hc <- hclust(dist(x))
plot(hc)

abline(h = 1.8, col = "red")
abline(h = 2.2, col = "blue")
```

Q. How does this compare to your known 'col' groups?
```{r}
gp2 <- cutree(hc, k = 2)
gp3 <- cutree(hc, k = 3)
```

```{r}
gp2
gp3
plot(x, col = gp3)
```

```{r}
table(gp2, gp3)
```

# Principal Component Analysis

We will use the **prcomp()** function for PCA
```{r}
# example data
mydata <- read.csv("https://tinyurl.com/expression-CSV",
row.names=1)
nrow(mydata)
ncol(mydata)
colnames(mydata)
pca <- prcomp(t(mydata), scale = TRUE)
attributes(pca)
```

```{r}
plot(pca$x[,1], pca$x[,2])
pca.var <- pca$sdev^2
pca.var
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
pca.var.per
```

```{r}
barplot(pca.var.per, main="Scree Plot",
xlab="Principal Component", ylab="Percent Variation")
```

```{r}
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
xlab=paste0("PC1 (", pca.var.per[1], "%)"),
ylab=paste0("PC2 (", pca.var.per[2], "%)"))
```

```{r}
identify(pca$x[,1], pca$x[,2], labels=colnames(mydata))
```

