---
title: "Class7"
author: "Sharon Chang"
date: "4/23/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#  Functions revisited

We will source a file from online with our functions from last time

```{r}
source("http://tinyurl.com/rescale-R")
```

Try out the last day's rescale function
```{r}
rescale(1:10)
```

Try the rescale2() function that catches string inputs
```{r}
rescale2(c(1:10), "string"))
```

# Find missing NA values in teo vectors

Start with a simple example of the larger problem we are trying to solve
```{r}
x <- c(1, 2, NA, 3, 4)
y <- c(NA, 3, NA, 3, 4)
```

```{r}
is.na(x)
```

```{r}
is.na(y)
```

Try putting these together with an AND
```{r}
is.na(x) & is.na(y)
```

Take the sum() to find out how many TRUE values we have and thus how many NAs we have in both x and y
```{r}
sum(is.na(x) & is.na(y))
```

Now we can make this into our first function
```{r}
both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
}
```

```{r}
both_na(x, c(NA, 3, NA, 2, NA))
```

Test
```{r}
x <- c(NA, NA, NA)
y1 <- c(1, NA, NA)
y2 <- c(1, NA, NA, NA)
```

```{r}
both_na(x, y2)
```

```{r}
y3 <- c(1, NA, NA, NA, NA, NA, NA)
both_na(x, y3)
```

```{r}
3 == 3
```

```{r}
3 < 2
```

```{r}
length(x)
```

```{r}
length(y2)
```

Now let's try both_na2() function on our different length input vectors
```{r}
both_na2(x, y2)
```

```{r}
which(c(F, F, T, F))
```

```{r}
which(is.na(c(1, 2, NA, 4)))
```

```{r}
x <- c(1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

both_na3(x, y)
```

#Intersect function
```{r}
df1
```

Make things simple
```{r}
x <- df1$IDs
y <- df2$IDs

x
y
```

```{r}
intersect(x, y)
```

```{r}
which(x %in% y)
```

```{r}
cbind( x[x %in% y],
       y[y %in% x])
```

Use the RStudio shortcut 'CODE > EXTRACT FUNCTION' to turn our snippet into a working function
```{r}
gene_intersect <- function(x, y) {
  cbind( x[x %in% y],
         y[y %in% x])
}
```

```{r}
gene_intersect(df1$IDs, df2$IDs)
```

```{r}
gene_intersect2(df1, df2)
```

```{r}
gene_intersect3(df1, df2)
```

```{r}
merge(df1, df2, by = "IDs")
```

Grade: to calculate mean score after dropping the lowest score
```{r}
x <- c(100, 90, 90, 90, 90, 90, 97, 80)

grade <- function(x) {
  (sum(x) - min(x))/ (length(x)- 1)
}

grade(x)
```

